\section{Federated Learning}\label{section:federated_learning}

This section contains necessary background information and context regarding FL.
The first subsection covers fundamental FL building blocks and terminologies.
The next subsection explains vital supplementary FL concepts.
Oakestra orchestrates FLOps.
It uses an unconventional three-tiered structure that allows support for geographical clusters \cite{paper:oakestra_usenix}.
This structure opens up unique opportunities for FL applications.
More advanced FL architectures are necessary to benefit from these opportunities.
The following subsection discusses these architectures.
These three subsections build a solid FL understanding.
A great summary and source of information for deeper insights into the field of FL can be found in \cite{book:fl}.
It is a 2022 book that captures and discusses the history and progress of FL research and state-of-the-art FL techniques.

Building on top of established FL understanding, the subsequent subsection reviews the research landscape of FL.
That subsection showcases active and popular research directions.
It points out underexplored aspects and weak points in the field.
The penultimate subsection analyzes and compares existing FL frameworks and libraries.
The concluding subsection provides an overview of the FL framework FLOps uses.

\input{chapters/background/sections/federated_learning/subsections/fl_basics.tex}
\input{chapters/background/sections/federated_learning/subsections/fl_supplementary_concepts.tex}
\input{chapters/background/sections/federated_learning/subsections/fl_architectures.tex}
\input{chapters/background/sections/federated_learning/subsections/fl_research.tex}
\input{chapters/background/sections/federated_learning/subsections/fl_frameworks_and_libraries.tex}
\input{chapters/background/sections/federated_learning/subsections/flower.tex}