\chapter{Background}\label{chapter:background}

As mentioned in the contributions section, FLOps combines and uses a large set of technologies from
different disciplines. To properly understand FLOps as a whole and why it combines these techniques,
it is necessary to analyze them individually.
This enables us to form a common understanding, including critical background knowledge of their benefits and downsides.
Only afterward does it make sense to discuss how FLOps merges them to create something new.

This background chapter provides a general overview for each sector and discusses aspects that
are necessary for FLOps in greater detail.
We start with exploring the field of federated learning.
FL is the core task at hand that FLOps aims to optimize.
A thorough understanding of this discipline is required to figure out where it has shortcomings.
To improve upon these weaknesses, we study the established set of 
best practices from DevOps and MLOps.
Techniques like automation and CI/CD require infrastructure and resources.
Orchestration allows us to provision, manage, and deploy such infrastructure and resources.
We review the field of orchestration technologies and provide a short
overview of Oakestra \cite{paper:oakestra_usenix} as the chosen platform.
In the final background section, we take a look at and compare a couple of existing pieces of work
that resemble FLOps.


\input{chapters/background/sections/federated_learning/section.tex}

\input{chapters/background/sections/mlops/section.tex}

\input{chapters/background/sections/orchestration.tex}

\input{chapters/background/sections/related_work.tex}